### 2024-10-22 manuscript.tex
Various improvements

Increasing climate model resolution is a common way of improving model accuracy
\citep{mauritsen2022}. This is of course limited by the available computational
power and a tradeoff with increasing parametrisation complexity. Current
computational availability and acceleration from general-purpose computing on
graphics processing units (GPGPUs) is progressing to enable km-scale (also
called k-scale) Earth system models (ESMs) and coupled atmosphsere--ocean
general circulation models (AOGCMs) in research conditions today and
operationally in the forthcoming years. Therefore, it represents a natural
advancement in climate modeling.  Global strom-resolving models (GSRMs) are
emerging as a new front in the development of high-resolution global climate
models, with horizontal grid resolutions of about 2--8 km
\citep{satoh2019,stevens2019}. This is enough to resolve mesoscale convective
storms, but smaller-scale convective plumes and smaller-scale cloud structure
remain unresolved. At a 5-km scale, non-hydrostatic processes also become
important, and for this reason such models are generally non-hydrostatic. The
terms global cloud-resolving models (GCRMs), or global
convection-permitting/-resolving models (GCPMs) are also sometimes used
interchangeably with GSRMs but imply that clouds or convection are resolved
explicitly, which is not entirely true for GSRMs, as this would require an even
higher horizontal resolution \citep{satoh2019}.  Representative of these
efforts is the DYnamics of the Atmospheric general circulation Modeled On
Non-hydrostatic Domains (DYAMOND) project \citep{stevens2019,dyamond}, which is
an intercomparison of nine global storm-resolving models (GSRMs) on two 40-day
time periods in summer (1 August -- 10 September 2016) and winter (20 January -
1 March 2020). A new one-year GSRM intercomparison is currently proposed by
\cite{takasuka2024}, with the hope of also evaluating the seasonal cycle and
large-scale circulation.  An alternative to using a computantionally costly
GSRM is to train an artificial neural network on GSRM output and use it for
subgrid-scale clouds, as done with the GSRM ICON by \cite{grundner2022};
\cite{grundner2023}.

@@

In general, SST biases in the SO can originate either in the atmosphere, caused
by too much shortwave heating of the surface, too little longwave cooling of
the surface, or in the ocean circulation.  Interactions of both are also
possible, due to, for example SST affecting clouds and clouds affecting the
surface radiation.  \cite{zhang2023} have shown that SST biases have improved
in CMIP6 compared to CMIP5 (relative to ERA5), with SST overall increasing in
the later CMIP phase. However, over the SO this resulted in an even higher
positive bias, especially in the Atlantic Ocean (AO) sector of the SO,
increasing by up to 1°C.  \cite{luo2023} identified that the SO SST bias in an
ensemble of 18 CMIP6 models originates not from the surface heat and radiation
fluxes (using reanalyses as a reference), but from a warm bias in the Northern
Atlantic Deep Water.

The main aim of this study is to evaluate the GSRM version of ICON, developed
jointly by nextGEMS, Deutscher Wetterdienst (DWD), Max-Planck-Institute for
Meteorology (MPI-M), Deutsches Klimarechenzentrum (DKRZ), Karlsruhe Institute
of Technology (KIT), and the Center for Climate Systems Modeling (C2SM).
Previous studies have identified substantial large-scale biases in climate
model clouds over the SO, affecting sea surface temperature and the Earth’s
albedo. Our aim is to quantify how well the GSRM ICON is simulating clouds in
this region, particularly in light of the fact that subgrid-scale clouds and
convection are not parameterised in this model. This region is mostly dominated
by boundary layer clouds generated by shallow convection, and these are
problematic to observe by spaceborne lidars and radars, which are affected by
attenuation by overlapping and thick clouds and ground clutter, respectively.
Specifically, the radar on CloudSat and lidar on CALIPSO (neither of them
operational any more) are affected by the abovementioned issues, resulting in a
strong underestimation of cloud occurrence below 2 km relative to ground-based
lidar observations \citep{mcerlich2021}.  This, in turn, can lead to systematic
biases in low clouds in climate models, which are frequently evaluated against
CloudSat--CALIPSO products. Reanalyses can also suffer from cloud biases, as
these are usually parametrised in their atmospheric component, and also in
regions where input observations are sparse.  This makes them a problematic
reference for clouds over the SO, and any biases relative to a reanalysis
should be interpreted with caution. Instead, we chose to use a large set of
ship-based observations conducted with ceilometers and lidars on board of the
RV \emph{Polarstern} and other voyages and stations as a reference for the
model evaluation.

@@

Missing days in the ceilometer data were HMNZSW16 (7 days): 24--27 November, 10
December, 16--17 December 2016; MARCUS (3 days): 8, 10 November, 10 December;
MICRE (9 days): 7--8, 29 June, 5, 16 July, 15 August, 17 October 2016, 11
February, 21 March 2017; TAN1502 (1 day): 24 January.

@@

The Vaisala CL51 (photo in Fig. \ref{fig:map}d) and CT25K are ceilometers
operating at a near-infrared wavelength of 910 nm and 905 nm, respectively.
CL51 can also be configured to emulate the Vaisala CL31. The maximum range is
15.4 km (CL51), 7.7 km (CL31 emulation mode with 5 m vertical resolution), and
7.5 km (CT25K). The vertical resolution is 10 m (5 m configurable) in CL51 and
30 m in CT25K. The sampling (temporal) resolution is configurable and in our
datasets it is about 6 s for CL51 on AA15‐16, 16 s for CT25K on MARCUS and
MICRE, 36 s for CL51 on RV \emph{Polarstern}, and about 2.37 s for CL51 with
CL31 emulation on TAN1502. The wavelength of 910 nm is affected by water vapour
absorption of about 20\% in the mid-latitudes \citep{wiegner2015,wiegner2019},
but we do not expect this to be a significant issue as explained in
\cite{kuma2021}.  The instrument data files containing raw uncalibrated
backscatter were first converted to NetCDF with cl2nc
(\url{https://github.com/peterkuma/cl2nc}) and then processed with the ALCF
(Section \ref{sec:alcf}) to produce absolutely calibrated attenuated volume
backscattering coefficinet (AVBC), cloud mask, cloud occurrence by height and
the total cloud fraction. Because the CT25K uses a very similar wavelength to
CL51, equivalent calculations as for CL51 were done assuming a wavelength of
910 nm. The Vaisala CL51 and CT25K instruments were used on most of the voyages
and stations analysed here. Fig.  \ref{fig:example}a shows an example of AVBC
derived from the CL51 instrument data.

@@

ERA5 \citep{era5} is a reanalysis produced by the European Centre for
Medium-Range Weather Forecasts (ECMWF). It is based on a numerical weather
prediction (NWP) model Integrated Forecast System (IFS) version CY41R2.  The
horizontal resolution is 0.25° in latitude and longitude, which is about 28 km
in the North--South direction and 14 km in the East--West direction at 60°S.
Internally, the model uses 137 vertical levels. Here, we use output at 1-hourly
instantaeous time intervals, except for radiation quantities, which are
accumulations (from these we calculate daily means).  Vertically-resolved
quantities are made available on 37 pressure levels.

@@

TOA radiation quantities are taken from the Clouds and the Earth’s Radiant
Energy System (CERES) instruments on board of the Terra and Aqua satellites
\citep{wielicki1996,loeb2018}. In our analysis we used the adjusted all sky SW
and LW upwelling fluxes at TOA from the synoptic TOA and surface fluxes and
clouds 1 degree daily edition 4A product
(CER\_SYN1deg-Day\_Terra-Aqua-MODIS\_Edition4A)
\citep{doelling2013,doelling2016}.

@@

Here, we briefly describe the results of the cyclonic activity and stability
distribution, which is relavant for the subsequent analaysis, because these
conditions are used for subsetting our dataset. Fig.
\ref{fig:cyclone-stability}a shows a geographical distribution of the fraction
of cyclonic days as determined by the cyclone tracking algorithm applied on the
ERA5 reanalysis and ICON data (Section \ref{sec:cyclone-stability}). As
expected, the strongest cyclonic activity is in the high-latitude SO zone, and
it is relatively zonally symmetric at all latitudes.  While both reanalysis and
the model agree relatively well, differences in the strength of the local
extremes of occurrence are notable, especially over the Amundsen Sea, which is
more cyclonic in the reanalysis, and around Cape Adere, which is more cyclonic
in ICON. These differences might, however, stem from the relatively short time
periods of comparison (4 year) and the fact that the model is free-running.

>>>>

Increasing climate model resolution is a common way of improving model accuracy
\citep{mauritsen2022}. This is of course limited by the available computational
power and a tradeoff with increasing parametrisation complexity. Current
computational availability and acceleration from general-purpose computing on
graphics processing units (GPGPUs) is progressing to enable km-scale (also
called k-scale) Earth system models (ESMs) and coupled atmosphsere--ocean
general circulation models (AOGCMs) in research conditions today and
operationally in the forthcoming years. Therefore, it represents a natural
advancement in climate modeling.  Global strom-resolving models (GSRMs) are
emerging as a new front in the development of high-resolution global climate
models, with horizontal grid resolutions of about 2--8 km
\citep{satoh2019,stevens2019}. This is enough to resolve mesoscale convective
storms, but smaller-scale convective plumes and smaller-scale cloud structure
remain unresolved. At a 5-km scale, non-hydrostatic processes also become
important, and for this reason such models are generally non-hydrostatic. The
terms global cloud-resolving models, or global convection-permitting/-resolving
models are also sometimes used interchangeably with GSRMs but imply that clouds
or convection are resolved explicitly, which is not entirely true for GSRMs, as
this would require an even higher horizontal resolution \citep{satoh2019}.
Representative of these efforts is the DYnamics of the Atmospheric general
circulation Modeled On Non-hydrostatic Domains (DYAMOND) project
\citep{stevens2019,dyamond}, which is an intercomparison of nine global
storm-resolving models (GSRMs) on two 40-day time periods in summer (1 August
-- 10 September 2016) and winter (20 January - 1 March 2020). A new one-year
GSRM intercomparison is currently proposed by \cite{takasuka2024}, with the
hope of also evaluating the seasonal cycle and large-scale circulation.  An
alternative to using a computantionally costly GSRM is to train an artificial
neural network on GSRM output and use it for subgrid-scale clouds, as done with
the GSRM ICON by \cite{grundner2022}; \cite{grundner2023}.

@@

In general, sea surface temperature (SST) biases in the SO can originate either
in the atmosphere, caused by too much shortwave heating of the surface, too
little longwave cooling of the surface, or in the ocean circulation.
Interactions of both are also possible, due to, for example SST affecting
clouds and clouds affecting the surface radiation.  \cite{zhang2023} have shown
that SST biases have improved in CMIP6 compared to CMIP5 (relative to ERA5),
with SST overall increasing in the later CMIP phase. However, over the SO this
resulted in an even higher positive bias, especially in the Atlantic Ocean (AO)
sector of the SO, increasing by up to 1°C.  \cite{luo2023} identified that the
SO SST bias in an ensemble of 18 CMIP6 models originates not from the surface
heat and radiation fluxes (using reanalyses as a reference), but from a warm
bias in the Northern Atlantic Deep Water.

The main aim of this study is to evaluate the GSRM version of ICON, developed
jointly by nextGEMS, Deutscher Wetterdienst, Max-Planck-Institute for
Meteorology, Deutsches Klimarechenzentrum (DKRZ), Karlsruhe Institute of
Technology, and the Center for Climate Systems Modeling.  Previous studies have
identified substantial large-scale biases in climate model clouds over the SO,
affecting sea surface temperature and the Earth’s albedo. Our aim is to
quantify how well the GSRM ICON is simulating clouds in this region,
particularly in light of the fact that subgrid-scale clouds and convection are
not parameterised in this model. This region is mostly dominated by boundary
layer clouds generated by shallow convection, and these are problematic to
observe by spaceborne lidars and radars, which are affected by attenuation by
overlapping and thick clouds and ground clutter, respectively.  Specifically,
the radar on CloudSat and lidar on CALIPSO (neither of them operational any
more) are affected by the abovementioned issues, resulting in a strong
underestimation of cloud occurrence below 2 km relative to ground-based lidar
observations \citep{mcerlich2021}.  This, in turn, can lead to systematic
biases in low clouds in climate models, which are frequently evaluated against
CloudSat--CALIPSO products. Reanalyses can also suffer from cloud biases, as
these are usually parametrised in their atmospheric component, and also in
regions where input observations are sparse.  This makes them a problematic
reference for clouds over the SO, and any biases relative to a reanalysis
should be interpreted with caution. Instead, we chose to use a large set of
ship-based observations conducted with ceilometers and lidars on board of the
RV \emph{Polarstern} and other voyages and stations as a reference for the
model evaluation.

@@

Missing days in the ceilometer data were HMNZSW16 (7 days): 24--27 November, 10
December, 16--17 December 2016; Measurements of Aerosols, Radiation, and CloUds
over the Southern Ocean (MARCUS;  3 days): 8, 10 November, 10 December;
Macquarie Island Cloud Radiation Experiment (MICRE; 9 days): 7--8, 29 June, 5,
16 July, 15 August, 17 October 2016, 11 February, 21 March 2017; TAN1502 (1
day): 24 January.

@@

The Vaisala CL51 (photo in Fig. \ref{fig:map}d) and CT25K are ceilometers
operating at a near-infrared wavelength of 910 nm and 905 nm, respectively.
CL51 can also be configured to emulate the Vaisala CL31. The maximum range is
15.4 km (CL51), 7.7 km (CL31 emulation mode with 5 m vertical resolution), and
7.5 km (CT25K). The vertical resolution is 10 m (5 m configurable) in CL51 and
30 m in CT25K. The sampling (temporal) resolution is configurable and in our
datasets it is about 6 s for CL51 on AA15‐16, 16 s for CT25K on MARCUS and
MICRE, 36 s for CL51 on RV \emph{Polarstern}, and about 2.37 s for CL51 with
CL31 emulation on TAN1502. The wavelength of 910 nm is affected by water vapour
absorption of about 20\% in the mid-latitudes \citep{wiegner2015,wiegner2019},
but we do not expect this to be a significant issue as explained in
\cite{kuma2021}.  The instrument data files containing raw uncalibrated
backscatter were first converted to Network Common Data Form (NetCDF) with
cl2nc (\url{https://github.com/peterkuma/cl2nc}) and then processed with the
ALCF (Section \ref{sec:alcf}) to produce absolutely calibrated attenuated
volume backscattering coefficinet (AVBC), cloud mask, cloud occurrence by
height and the total cloud fraction. Because the CT25K uses a very similar
wavelength to CL51, equivalent calculations as for CL51 were done assuming a
wavelength of 910 nm. The Vaisala CL51 and CT25K instruments were used on most
of the voyages and stations analysed here. Fig.  \ref{fig:example}a shows an
example of AVBC derived from the CL51 instrument data.

@@

ERA5 \citep{era5} is a reanalysis produced by the European Centre for
Medium-Range Weather Forecasts (ECMWF). It is based on a numerical weather
prediction model IFS version CY41R2.  The horizontal resolution is 0.25° in
latitude and longitude, which is about 28 km in the North--South direction and
14 km in the East--West direction at 60°S.  Internally, the model uses 137
vertical levels. Here, we use output at 1-hourly instantaeous time intervals,
except for radiation quantities, which are accumulations (from these we
calculate daily means).  Vertically-resolved quantities are made available on
37 pressure levels.

@@

TOA radiation quantities are taken from the CERES instruments on board of the
Terra and Aqua satellites \citep{wielicki1996,loeb2018}. In our analysis we
used the adjusted all sky SW and LW upwelling fluxes at TOA from the synoptic
TOA and surface fluxes and clouds 1 degree daily edition 4A product
(CER\_SYN1deg-Day\_Terra-Aqua-MODIS\_Edition4A)
\citep{doelling2013,doelling2016}.

@@

Here, we briefly describe the results of the cyclonic activity and stability
distribution, which is relavant for the subsequent analaysis, because these
conditions are used for subsetting our dataset. Fig.
\ref{fig:cyclone-stability}a shows a geographical distribution of the fraction
of cyclonic days as determined by the cyclone tracking algorithm applied on the
ERA5 reanalysis and ICON data (Section \ref{sec:cyclone-stability}). As
expected, the strongest cyclonic activity is in the high-latitude SO zone, and
it is relatively zonally symmetric at all latitudes.  While both reanalysis and
the model agree relatively well, differences in the strength of the local
extremes of occurrence are notable, especially over the Amundsen Sea, which is
more cyclonic in the reanalysis, and around Cape Adare, which is more cyclonic
in ICON. These differences might, however, stem from the relatively short time
periods of comparison (4 year) and the fact that the model is free-running.
